Is high social class always beneficial for survival? Northern Sweden 1801--2013

### 2018-03-29 article3c.Rmd
fix description of presence periods

Differences in available information in the two datasets as well as in
the Linnaeus data make it necessary to apply different approaches when it
comes to the identification of presence periods. The Poplink data provide us with exact dates, or at least year of start and exit of presence, allowing us
to have full and continuous control over the \emph{de jure} population. This is
not the case with the Linnaeus database, however. For the period 1976--2001
we use presence in the censuses 1975--1990 and information on deaths from the National
Board of Health and Welfare. The populations in the Skellefteå and Umeå regions are those residing there according to the censuses. Each census constitute the baseline for which persons are followed up during the five years until the next census (e.g. the 1975 census has a follow-up from 1 January 1976 until the end of 1980). The exception of this is the 1990 census for which the follow-up is until 2001. Thus, the deaths do not necessarily take place in our region for the Linnaeus analyses, although this is usually the case.

For the last period, 2002--2013, we use the yearly population registers
(LISA data) together with information on deaths from the National Board of
Health and Welfare. This allows for full control over the presence periods
in the studied regions.

>>>>

Differences in available information in the two datasets as well as in
the Linnaeus data make it necessary to apply different approaches when it
comes to the identification of presence periods. The Poplink data provide us with exact dates, or at least year of start and exit of presence, allowing us
to have full and continuous control over the \emph{de jure} population. This is
not the case with the Linnaeus database, however. For the period 1976--2001
we use presence in the censuses 1975--1990 and information on deaths from the National
Board of Health and Welfare. The populations in the Skellefteå and Umeå regions are those residing there according to the censuses. Each census constitute the baseline for which persons are followed up during the five years until the next census (e.g. the 1975 census has a follow-up from 1 November 1975 until 31 October 1980). The exception of this is the 1990 census for which the follow-up is until 31 December 2001. Thus, the deaths do not necessarily take place in our region for the Linnaeus analyses, although this is usually the case.

For the last period, from 1 January 2002 until 31 December 2013, we use the yearly population registers
(LISA data) together with information on deaths from the National Board of
Health and Welfare. The LISA data depict the situation at the end of the given year, so the follow-up period is from 1 January to 31 December the next year. Deaths may still occur outside the region, but less frequently than in the Fob case.

### 2018-03-29 marginal.Rmd
add timing (temporarily)
fix description of presence periods

---
title: "Is high social class always beneficial for survival? Northern Sweden 1801--2013"
author: Sören Edvinsson\footnote{Corresponding author; Phone number +46907866040, email soren.edvinsson\@umu.se} and Göran Broström
date: 'Centre for Demographic and Ageing Research \newline Umeå University, SE-90187 Umeå, Sweden'
output:
  bookdown::pdf_document2:
    citation_package: natbib
    number_sections: yes
    toc: no
    toc_depth: 2
    keep_tex: true
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: yes
    toc_depth: 2
titlepage: yes
bibliography: "mybib.bib"
documentclass: article
classoption: titlepage
fontsize: 12pt
citation_package: natbib
biblio-style: "apalike"
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage{graphicx}
abstract: "Focusing on two regions in northern Sweden 1801--2013, we challenge common notions of the assumed advantage in survival of belonging to a high social class. The issue is analysed according to gender and age group (adults and elderly) and in relation to the development of economic inequality. The results show that high social class is not always favourable for survival. Men in the elite category, particularly in working age, had higher mortality compared to others during a large part of the studied period; a male mortality class reversal appears at a surprisingly late date, while the social gradient among women conforms to the expected pattern. We suggest that health-related behaviour is decisive not only in later but earlier phases of the mortality transition as well. The results implicate that the association between social class and health is more complex than is assumed in many of the dominant theories in demography and epidemiology."
---

@@

>>>>

---
title: "Is high social class always beneficial for survival? Northern Sweden 1801--2013"
author: Sören Edvinsson\footnote{Corresponding author; Phone number +46907866040, email soren.edvinsson@umu.se} and Göran Broström
date: 'Centre for Demographic and Ageing Research \newline Umeå University, SE-90187 Umeå, Sweden `r Sys.time()`'
output:
  bookdown::pdf_document2:
    citation_package: natbib
    number_sections: yes
    toc: no
    toc_depth: 2
    keep_tex: true
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: yes
    toc_depth: 2
titlepage: yes
bibliography: "mybib.bib"
documentclass: article
classoption: titlepage
fontsize: 12pt
citation_package: natbib
biblio-style: "apalike"
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage{graphicx}
abstract: "Focusing on two regions in northern Sweden 1801--2013, we challenge common notions of the assumed advantage in survival of belonging to a high social class. The issue is analysed according to gender and age group (adults and elderly) and in relation to the development of economic inequality. The results show that high social class is not always favourable for survival. Men in the elite category, particularly in working age, had higher mortality compared to others during a large part of the studied period; a male mortality class reversal appears at a surprisingly late date, while the social gradient among women conforms to the expected pattern. We suggest that health-related behaviour is decisive not only in later but earlier phases of the mortality transition as well. The results implicate that the association between social class and health is more complex than is assumed in many of the dominant theories in demography and epidemiology."
---

```{r time1, echo = FALSE}
time1 <- Sys.time() # Time the document rendering
```

@@

```{r time2, echo = FALSE}
time2 <- Sys.time()
time2 - time1
```
